Fastest Paths Between Points in $\R^2$ for a Particle Moving with Constant Speed and Bounded Centripetal Acceleration

### 2013-10-11 fixed_speed_theorems.tex
More changes to minimizing theta theorem.

\begin{lemma}\label{lem:too-close}
  Let particle $p$ have constant speed, bounded centripetal acceleration, an initial angle $\theta_0$, starting position $\bvec{X_0}$, and ending position $\bvec{X_f}$. Every valid path where $\theta(t)$ is minimized for all $t \in [T_0(\bhat{\gamma_p}), T_f(\bhat{\gamma_p})]$ has $d(\bvec{X}_0, \bvec{X}_f) \ge 2 R_{min} \cos\paren{\theta_0}$.
\end{lemma}

@@

  \includefig{valid_trajs.eps}{Valid and invalid final positions for a particle starting at position $X_0 = A$}{valid-traj}

  Looking at Figure \ref{fig:valid-traj}, one should notice that if $X_f$ is any position inside the circle of radius $R_{min}$, such as point B, then the path is invalid. It is not possible to turn inside the circle, because the centripetal acceleration is bounded. Thus, C is the closest valid ending position. C is a distance $2 R_{min} \cos\paren{\theta_0}$ away from A.

@@

\begin{conjecture}\label{conj:optimal-position}
  Let $p_1, p_2$ be particles with constant speed, bounded centripetal acceleration, starting positions $\bvec{X}_1, \bvec{X}_2$ respectively, and a final position $\bvec{X}_f$, such that $d(\bvec{X}_1, \bvec{X}_f), d(\bvec{X}_2, \bvec{X}_f) \ge 2 R_{min} \cos\paren{\theta_0}$. If $r_1(T_0) \leq r_2(T_0)$ and $\abs{\theta_1(T_0)} < \abs{\theta_1(T_0)}$, then $T_f(\hat{\gamma_1}) < T_f(\hat{\gamma_2})$.
\end{conjecture}

@@

The special case when $d(\bvec{X}_1, \bvec{X}_f), d(\bvec{X}_2, \bvec{X}_f) < (1/2) R_{min} \cos\paren{\theta_0}$ is omitted, because some special strategy will be needed in this situation. We leave this case as an open problem in our paper.

@@

We know that $\abs{a_{c,1}(t)} = \bar{a_c}$ for all $t < t_s$ because $p_1$ follows a path that minimizes $\abs{\theta_1(t)}$. Thus, we see that $p_1$ traces out an arc of a circle with radius $R_{min} = \frac{\bar{v}}{\bar{a}_c}$.

Now let us introduce a particle $p_2$ with the same conditions as $p_1$, except that at some time $t_b < t_s$, $p_2$ will choose $a_{c,2}$ such that $\abs{a_{c,2}} \neq \bar{a}_c$. We shall show that $\hat{\gamma_2(t)}$ after this divergence is strictly slower than $\hat{\gamma_1}$, which will allow us to conclude that $\gamma_1(t)$ is the unique fastest path.

@@

\begin{align*}
  \frac{d^2 r_1(t)}{d t^2} - \frac{d^2 r_2(t)}{d t^2} = - \sin\paren{\abs{\theta(t_b)}} \paren{\bar{a}_c - a_{c,2}(t_b)}
  & < 0
\end{align*}

The inequality was obtained because $\bar{a}_c - a^{p_2}_c(t_b) > 0$. This directly implies that $r_1(t_b + \epsilon) < r_2(t_b + \epsilon)$ for some small $\epsilon > 0$. Therefore the distance left to travel for $p_1$ is less than the distance left for $p_2$ at $t_b + \epsilon$ time, i.e. $d(\gamma_1(t_b + \epsilon), \bvec{X}_f) < d(\gamma_2(t_b + \epsilon), \bvec{X}_f)$.

We can also use equation \ref{eq:theta-diff} to obtain the following for the difference in angles:
\begin{eqnarray}
  \frac{d \abs{\theta_1(t_b)}}{dt} - \frac{\abs{\theta_2 (t_b)}}{dt} = - \frac{1}{v(t_b)} \left(\bar{a}_c - a^{p_2}_c(t_b) \right) < 0
\end{eqnarray}

Again, we used the fact that $\bar{a}_c - a^{p_2}_c(t_b) > 0$ to show the inequality. This implies that $\theta_1(t_b + \epsilon) < \theta_2(t_b + \epsilon)$.

Now, we have satisfied the conditions of Conjecture \ref{conj:optimal-position} since $d(\gamma_1(t_b + \epsilon), \bvec{X}_f) < d(\gamma_2(t_b + \epsilon), \bvec{X}_f)$ and $\theta_1(t_b + \epsilon) < \theta_2(t_b + \epsilon)$. This implies that $T_f(\gamma_1) < T_f(\gamma_2)$. Thus, the fastest path before $t_s$ must be to travel along $\gamma_1$. However, we know that the fastest path after $t_s$ is to travel along the straight line (which $\gamma_1$ does). Therefore, we see that $\gamma_1 = \bhat{\gamma}_p$ is the fastest path.
\end{proof}

>>>>

\begin{lemma}\label{lem:too-close}
  Let particle $p$ have constant speed, bounded centripetal acceleration, an initial angle $\theta_0$, starting position $\bvec{X_0}$, and ending position $\bvec{X_f}$. Every valid path where $\theta(t)$ is minimized for all $t \in [T_0(\bhat{\gamma_p}), T_f(\bhat{\gamma_p})]$ has $d(\bvec{X_0}, \bvec{X_f}) \ge 2 R_{min} \cos\paren{\theta_0}$.
\end{lemma}

@@

  \begin{figure}[H]
    \begin{center}
      \includegraphics[width=3.5in,height=2.5in,keepaspectratio]{valid_trajs.eps}
    \end{center}
  \vspace{-.2in} % corrects bad spacing
  \caption{\label{fig:valid-traj}}
  \end{figure}

  One should notice that if $X_f$ is any position inside the circle of radius $R_{min}$, such as point B, then the path is invalid. It is not possible to turn inside the circle, because the centripetal acceleration is bounded. Thus, C is the closest valid ending position. C is a distance $2 R_{min} \cos\paren{\theta_0}$ away from A.

@@

\begin{conjecture}\label{conj:optimal-position}
  Let $p_1, p_2$ be particles with constant speed, bounded centripetal acceleration, starting positions $\bvec{X_1}, \bvec{X_2}$ respectively, and a final position $\bvec{X}_f$, such that $d(\bvec{X_1}, \bvec{X}_f), d(\bvec{X_2}, \bvec{X}_f) \ge 2 R_{min} \cos\paren{\theta_0}$. If $r_1(T_0) \leq r_2(T_0)$ and $\abs{\theta_1(T_0)} < \abs{\theta_1(T_0)}$, then $T_f(\hat{\gamma_1}) < T_f(\hat{\gamma_2})$.
\end{conjecture}

@@

The special case when $d(\bvec{X_1}, \bvec{X}_f), d(\bvec{X_2}, \bvec{X}_f) < (1/2) R_{min} \cos\paren{\theta_0}$ is omitted, because some special strategy will be needed in this situation. We leave this case as an open problem in our paper.

@@

We know that $\abs{a_{c,1}(t)} = \bar{a}_c$ for all $t < t_s$ because $p_1$ follows a path that minimizes $\abs{\theta_1(t)}$.

Now let us introduce a particle $p_2$ with the same conditions as $p_1$, except that at some time $t_b < t_s$, $p_2$ will have $a_{c,2}(t) \neq \bar{a}_c$. We shall show that the fastest path for $p_2$, $\hat{\gamma_2(t)}$, after this divergence is strictly slower than the fastest path for $p_1$, $\hat{\gamma}_1$, which will allow us to conclude that $\gamma_1(t)$ is the unique fastest path.

@@

\begin{align*}
  \frac{d^2 r_1(t)}{d t^2} - \frac{d^2 r_2(t)}{d t^2}& = - \sin\paren{\abs{\theta(t_b)}} \paren{\bar{a}_c - a_{c,2}(t_b)}\\
  & < 0
\end{align*}

The inequality was obtained because $\bar{a}_c - a_{c,2}(t_b) > 0$. This directly implies that $r_1(t_b + \epsilon) < r_2(t_b + \epsilon)$ for some small $\epsilon > 0$. Therefore the distance left to travel for $p_1$ is less than the distance left for $p_2$ at $t_b + \epsilon$ time, i.e. $d(\gamma_1(t_b + \epsilon), \bvec{X_f}) < d(\gamma_2(t_b + \epsilon), \bvec{X_f})$.

We can also use equation \ref{eq:theta-diff} to obtain the following for the difference in angles:
\begin{align*}
  \frac{d \abs{\theta_1(t_b)}}{dt} - \frac{\abs{\theta_2 (t_b)}}{dt}& = - \frac{1}{v(t_b)} \paren{\bar{a}_c - a_{c,2}(t_b)}\\
  & < 0
\end{align*}

Again, we used the fact that $\bar{a}_c - a_{c,2}(t_b) > 0$ to show the inequality. This implies that $\abs{\theta_1(t_b + \epsilon)} < \abs{\theta_2(t_b + \epsilon)}$.

Now, we have satisfied the conditions of Conjecture \ref{conj:optimal-position} since $d(\gamma_1(t_b + \epsilon), \bvec{X_f}) < d(\gamma_2(t_b + \epsilon), \bvec{X_f})$ and $\abs{\theta_1(t_b + \epsilon)} < \abs{\theta_2(t_b + \epsilon)}$. This implies that $T_f(\gamma_1) < T_f(\gamma_2)$. Thus, the fastest path before $t_s$ must be to travel along $\gamma_1$. However, we know that the fastest path after $t_s$ is to travel along the straight line (which $\gamma_1$ does). Therefore, we see that $\gamma_1$ is the unique fastest path.
\end{proof}

### 2013-10-11 particle_definitions.tex
More changes to minimizing theta theorem.

\includefig{polar_param.eps}{A particle moving in a B centered particle coordinate system}{polar-param}

@@

To find $\frac{d\beta(t)}{dt}$, we can look at a point, $p$, subject to nonzero centripetal acceleration, and zero tangential acceleration. The change in $\bvec{v(t)}$ over an infinitesimal time, $dt$, is shown in Figure \ref{fig:theta-deriv} (the two vectors, $\bvec{v(t)}$ and $\paren{\bvec{v(t)}+\bvec{dv(t)}}$, are superimposed).

\includefig{theta_deriv.eps}{Tangential acceleration}{theta-deriv}

>>>>

\begin{figure}[H]
    \begin{center}
      \includegraphics[width=4in,height=3in,keepaspectratio]{polar_param.eps}
    \end{center}
  \vspace{-.2in} % corrects bad spacing
  \caption{A particle moving in a B centered particle coordinate system.\label{fig:polar-param}}
  \end{figure}

@@

To find $\frac{d\beta(t)}{dt}$, we can look at a point, $p$, subject to nonzero centripetal acceleration, and zero tangential acceleration. The change in $\bvec{v(t)}$ over an infinitesimal time, $dt$, is shown below (the two vectors, $\bvec{v(t)}$ and $\paren{\bvec{v(t)}+\bvec{dv(t)}}$, are superimposed).

\begin{figure}[H]
    \begin{center}
      \includegraphics[width=3.5in,height=2.5in,keepaspectratio]{theta_deriv.eps}
    \end{center}
  \vspace{-.2in} % corrects bad spacing
  \caption{}
\end{figure}
